\chapter*{Useful Texts \& Further Readings}

\section*{Textbooks}
\begin{itemize}
    \item M. E. Peskin and D. V. Schroeder, “\textit{An Introduction to QFT}” (Addison Wesley, 1995).
    \item C. Itzykson and J.-B. Zuber, “\textit{Quantum Field Theory}” (McGraw-Hill,1980).
    \item L. H. Ryder, “\textit{Quantum Field Theory}” (Cambridge University Press, 1996). 
    \item P. Ramond, “\textit{Field Theory, A Modern Primer}” (Benjamin, 1994).
    \item S. Weinberg, “\textit{Quantum Theory of Fields}”, vol I and II (Cambridge University Press, 1996).
    \item A. Zee, “\textit{Quantum Field Theory in a Nutshell}” (Princeton University Press, 2010).
    \item F. Mandl and G. Shaw, “\textit{Quantum Field Theory}” (Wiley, 1984.
    \item L. H. Ryder, “\textit{Quantum Field Theory}” (Cambridge University Press, 1996).
    \item Schwartz, Matthew D. \textit{Quantum field theory and the standard model} (Cambridge University Press, 2014).
\end{itemize}

\section*{Other Similar Courses Available Online}
\begin{itemize}
    \item Prof. David Tong,  "\href{http://www.damtp.cam.ac.uk/user/tong/qft.html}{Lectures on Quantum Field Theory}", Cambridge University. 
    \item Prof. Michael Luke, "\href{https://www.physics.utoronto.ca/~luke/PHY2403F/References_files/lecturenotes.pdf}{PHY2403F Lecture Notes}", University Of Toronto. 
    \item Prof. Timo Weigand "\href{https://www.thphys.uni-heidelberg.de/~weigand/QFT2-14/SkriptQFT2.pdf}{Quantum Field Theory I + II}", Institute for Theoretical Physics, Heidelberg University. 
\end{itemize}